---
title: "Sample on Movie Ratings"
author: "John Cruz"
date: "2023-02-02"
output: pdf_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

I will be collecting movie ratings through an online survey on a scale of 1 to 5. 

The overall steps of this project is to do the following:

1. Collect survey data
2. Create a database to store this information
3. Able to query the data into R
4. Analyze some descriptive measures.

---

# Collecting Survey Data

The survey was conducted using a Google Forms and named *Spoiled Corn*. This survey will collect ratings of six recent popular movies on a scale of 1 to 5, 1 being the lowest. If a person did not watch this movie, it will be categorized with an option of N/A.

[Online Survey](https://forms.gle/Jf8fvruorDacr69j9)

# Building our Database

---

# Connecting to our Database

## Required Libraries

```{r libraries, warning=FALSE, message=FALSE}
library('RPostgres')
library('DBI')
library('tidyverse')
```

## PostgreSQL Database

Using this connection we will be able to keep our username and password from being embedded into our code and instead have RStudio pop up a dialog box to input the information. 

```{r DB-connection}
cnxn <- dbConnect(RPostgres::Postgres(),
                     dbname = 'movie_survey',
                     port = 5432,
                     user = 'postgres',
                     password = 'elliebelly968')
                     # user = rstudioapi::askForPassword("Database Username"),
                     # password = rstudioapi::askForPassword("Database Password"))
```

## Query the Database

Here we are able to directly query our PostgreSQL database and pull the information from the tables we created earlier. *dbFectch()* will allow us to move the results of the query into a readable DataFrame.

```{r DB-query}
query <- dbSendQuery(cnxn, 
                   "SELECT
                    	r.datetimeid,
                    	r.user_id,
                    	f.name,
                    	r.rating,
                    	f.release_year
                    FROM ratings r
                    INNER JOIN films f
                    	ON r.movie_id = f.movie_id ")

df <- dbFetch(query)
```

## Close Connection

Since we have gathered the data we needed from the database, we will close the connection to PostgreSQL.

```{r DB-disconect}
# Disconnect from database
dbDisconnect(cnxn)
```

